Autonomous Oceanographic Data Collection towards Estimating Excursion Sets of Vector-valued Random Fields

### 2020-06-18 main_new.tex
Updates from Overleaf

The remainder of this paper is organized as follows: 
%Section \ref{sec:bg} provides select background on ocean sampling. 
Section \ref{sec:ESEP} defines ESs, EPs, and the design criteria of IBV for
vector-valued GPs. Section \ref{sec:heuristics} builds on these
assumptions when deriving the sequential design criteria for adaptive
sampling. Section \ref{sec:simulations} discusses properties of the
methods in simulation studies. Section \ref{sec:case_study}
demonstrates the methodology used in field work characterizing a river
plume and finally, Section \ref{sec:concl_disc} contains a summary and
a discussion of future work.

%\subsection{}

%\begin{itemize}
%    \item 
%\end{itemize}

\begin{comment}
\section{Former Introduction (Before 08.06.2020)}

Monitoring the world's oceans has gained increased importance in light
of the changing climate and increasing anthropogenic impact. 
A central problem for understanding these factors is the lack of representative data with sufficient resolution. Most of this \emph{undersampling} can be attributed to the large spatio-temporal variations in which ocean processes transpire, prompting the need for effective means of sampling. By \emph{sampling}, we refer to the design of observational strategies in the spatial domain, where the use autonomous robotic platforms can be combined with statistical methods to pursue measurements with high scientific relevance. This combination, of statistical tools and robotic platforms, constitute a methodological basis for addressing the problem of efficient sampling of the ocean, which is at the core of this work.

The
advent of robust mobile robotic platforms \citep{Bellingham07} has
resulted in significant contributions to environmental monitoring and
sampling in the ocean. In particular, autonomous underwater vehicles (AUVs), have
advanced the state of sampling and consequently have made robotics an
integral part of ocean observation; our previous work has contributed
to this effort \citep{das11b,Das2015,fossuminformation,fossum18b}.

The potential that lies in using information-based sampling methods, which we
highlight in this paper, is to do efficient environmental sampling and
overcome common challenges as discussed next. First, full numerical ocean models based on complex
differential equations cannot be run onboard a marine robot with
limited computational capacity.
% provide accurate results for online
% operations because the onboard computers lack the computational power
% required.
Hence statistical proxy models of the environment must be used. For our purpose, we rely on Gaussian process (GP)
representations of the ocean variables of interest because they are computationally convenient and yet provide enough flexibility to realistically model the spatial variability and dependence, and the correlation between multiple processes. Second, because of limited resources to sample the large oceanographic domains, one must plan for
active learning during the operation, where a relevant criterion is
used to extract the most valuable designs. The acquired in-situ
information must then be assimilated into the statistical model
onboard the platform, so that it can be used to re-compute the
sampling criterion and in this way inform decisions on where to sample
next. We use a criterion focusing on the prediction of excursion sets
(ESs) for the multivariate spatial processes of interest. Third, the
number of sampling designs is enormous, creating a trade-off between
optimization and computability. One must often compromise in both
development and practical operations, and we will rely on
short-horizon optimization routines for sequential designs among
nearest neighbors points in a spatial grid (waypoint graph).

\kc{The para above needs a rewrite to me. It starts off by stating
  there are challenges. But then jumps to models without stating the
  connection with sampling, and then jumps into learning without
  stating why learning is important. Typically the English in formal
  text starts with a statement which is supported or clarified. In the
  two cases above, the connections are simply not obvious. If this was
  there in the first version of the m/s submitted, I'm not sure why
  the reviewers didn't catch it. This is not just a stylistic issue;
  it has to do with making sense and bring the reader/reviewer along.}

\kc{Second, I don't see a motivation for GPs here and as I read
  further into the text. Even for stats folks, are GPs the ``best''
  ways to go? Are they obviously the tools one would use in
  environmental sampling? And even if they are, I think we ought to
  make the case for it strongly and generally, so this paper is
  actually accessible and understandable to those not in the Spatial
  Stats field. I would suggest adding a short para on GPs and why
  they're important -- i.e. motivate the use of GPs early, so we can
  freely make attributions to them later, as in the next section, for
  example.}

The imperfect understanding of causality coupled with the underlying
fundamental uncertainty of the ocean implies that we need to make a
large number of measurements. In addition, the use statistical models
and methods to determine what and how phenomena form, is one way to
increase the quality of predictive power aiding our understanding of
upper water-column processes.

With this in mind, we design AUV sampling strategies for reducing the spatially-integrated Bernoulli variance (IBV) in the ES
indicator variables. 
ES of random fields are thoroughly presented in \citep{adler2009random}. Recent work on ESs and associated excursion
probabilities (EPs) connected to spatial statistics include
\cite{french2013spatio,bolin2015excursion,french2016credible}. Our
focus is on uncertainty characterization and how this is influenced by
sampling. In this sense, our work is similar to
\cite{bect2012,chevalier2014fast,azzimonti2016quantifying} who
describe analytical results for the IBV for univariate processes.

The remainder of this paper is organized as follows: Section
\ref{sec:bg} provides select background on ocean sampling. Section
\ref{sec:ESEP} defines ESs, EPs, and the design criteria of IBV for
vector-valued GPs. Section \ref{sec:heuristics} builds on these
assumptions when deriving the sequential design criteria for adaptive
sampling. Section \ref{sec:simulations} discusses properties of the
methods in simulation studies. Section \ref{sec:case_study}
demonstrates the methodology used in field work characterizing a river
plume and finally, Section \ref{sec:concl_disc} contains a summary and
a discussion of future work.
\end{comment}

\begin{comment}
\section{Ocean Processes and Observation}
\label{sec:bg}

Central to understanding the changes taking place in the upper
water-column is knowledge of the bio-geophysical interaction driven by
an agglomeration of physical forcings (e.g. wind, topography,
bathymetry, tidal influences, etc.) and incipient micro-biology driven
by planktonic and coastal anthropogenic input, such as pollution and
agricultural runoff transported into the ocean by rivers and streams.
These often result in a range of ecosystem-related phenomena such as
blooms and plumes, with direct and indirect effects on society. \kc{we
  should look for a citation here. Check with John Ryan}

Traditional data collection to understand such phenomena at sea has
typically been based on static buoys, floats, or ship-based methods,
with significant logistical limitations that directly impact coverage
and sampling resolution. Modern methods using satellite remote-sensing
provide large-scale coverage but have limited resolution, are limited
to sensing the surface, and are impacted by cloud cover. Numerical
ocean models similarly find it challenging to provide detail at fine
scale \citep{Lermusiaux:2006}. Mobile robotic platforms
\citep{Bellingham07} have contributed significantly to environmental monitoring and sampling. In particular, AUVs have
advanced the state of in-situ sampling, and made robotics an integral
part of ocean observation, filling parts of the undersampling gap
\cite{rudnick03,rudnick18}(See Fig. \ref{fig:envir1}).

%Walter Munk referred to the 20\textsuperscript{th} century
%as the \emph{century of undersampling} \citep{munk2002}, pointing out
%a lack of sufficient sampling resolution in time and space as a
%critical issue for characterizing ocean phenomena. Even today,
%policymakers struggle with limited observability of severe
%environmental problems; a prominent example is the recent (May 2018)
%outbreak of harmful algal blooms in Northern Norway\footnote{Report
%  from the Institute of Marine Research: https://bit.ly/2KZGJfI}, with
%losses exceeding 12 million tons of salmon valued at \$80 million
%(Norwegian Directorate of Fisheries). This is a case in point of how
%adequate sampling efforts could have provided a more accurate
%assessment and early warning of changing environmental conditions.

\begin{figure}[!h] 
  \centering 
  \subfigure[Illustration of an ocean sensing network and the sense-plan-act control methodology.]{\includegraphics[width =
    0.49\textwidth]{Figures/envir.pdf}\label{fig:envir1}}
  \hfill
  \subfigure[Frontal patterns off of the Nidelva river, Trondheim, Norway.]{\includegraphics[width =
    0.49\textwidth]{Figures/river_proccess.pdf}\label{fig:nidelven}}
  \caption{\ref{fig:envir1} Traditional ocean observation based on 
    ship-based sampling has been augmented by autonomous
    robotic vehicles. % and their interactions.  
    AUV platforms are an integral part of this network being able to
    reason and make decisions for efficient onboard adaptive sampling.
    % using the sense-plan-act control approach to autonomous control. 
    \ref{fig:nidelven} The interaction of river and ocean creates
    processes that are challenging to map, where the combination of
    statistics and robotics can play a vital role in enabling more
    effective oceanographic observation.\kc{The inset ``AUV robot''
      looks very tacky and ad-hoc; might be best to remove it. Change
      ``ships'' to
      ``Ship'', remove the aerial drone, since we didn't use it and
      don't plan to in this work.}}
  \label{fig:envir} \end{figure}

\end{comment}

\begin{comment}
Surveys with AUVs are usually limited to observations along fixed
transects that are pre-scripted in mission plans created manually by a
human operator. The mission can be specified operating on a scale of
hundreds of meters to tens of kilometers depending on the scientific
context. Faced with limited coverage capacity, a more effective
approach is to instead use onboard algorithms to continuously
evaluate, update, and refine future sampling locations (sense-plan-act
cycle in Fig. \ref{fig:envir1}), making the information gathering
\emph{adaptive} \citep{das11b,Das2015,fossuminformation,fossum18b}.
% This usually occurs on a spatial grid, called a waypoint graph. 
In doing so, the space of sampling opportunities is still limited by
the waypoint graph, but the AUV has flexibility to modify its path at
each waypoint based on in-situ sensing and measurements onboard
\citep{py10,Rajan12,Rajan12b}.

\end{comment}

\input{section2}

\begin{comment}
\section{Quantifying uncertainty on Excursion Sets implicitly defined by Gaussian processes}
\label{sec:ESEP}

We define the random sets of interest in Section \ref{sec:ES}, while the closed-form expressions for the expected uncertainty reduction are derived in Section \ref{sec:EIBV}. Section \ref{Sec:UnivarEx} provides an instructive bivariate example relevant for sampling in the temperature and salinity application.

\subsection{Excursion sets and Integrated Bernoulli variance}
\label{sec:ES}

Our objective is to leverage statistical tools onboard an autonomous
robotic platform to characterize a river plume, focusing on spatial
separation of cold freshwater from a river and warmer saline waters
of a fjord. The characterization of uncertainties on the ES is a
useful goal of this sampling strategy.

In order to formalize the problem in mathematical terms, let us
consider a multivariate (i.e., vector-valued) Gaussian random field
$(\gp[\x])_{\x \in \mathcal{M}}$, where $\mathcal{M}$ is an arbitrary
set (typically in geostatistical studies, a geographical subset of interest) and the
dimensionality of the output $\no \geq 1$ is arbitrary. We have $\no = 2$ in the example with temperature and salinity in the river plume.

Assume furthermore that the set of interest
is characterized in terms of a subset of values of the variables, say
$\T \subset \R^{\no}$, and denote the corresponding set of index
points by
$$
\es=\gp^{-1}(\T)=\{\x \in \mathcal{M}: \gp[\x] \in \T\}.
$$
Our goal here is to develop approaches to quantify and improve the
characterization of uncertainties on such sets $\es$, in the
context of multivariate random fields.

While some aspects of the developed approaches do not call for a
specific form of $\T$, we will often stick for simplicity to the case
of orthants
($\T=(-\infty, t_1] \times \dots \times (-\infty, t_{\no}]$ where
$t_1,\dots, t_{\no} \in \R$) as this will allow efficient calculation
of several key quantities. Note that changing some $\leq$ inequalities
to $\geq$ ones would lead to immediate adaptations. In the application
specific properties following the orthant assumption come into play
\kc{This last phrase doesn't parse; is it ``..., the following
  orthant....''?}: We let $t_1$ be a threshold for temperature and
$t_2$ a threshold for salinity \kc{Might be easier to remember with
  symbols 't' and 's' in some
  form.}. The set of index point then defines
$\mbox{ES} = \{\x \in D : \gp[\x,1] \leq t_1,\gp[\x,2] \leq t_2\}$.
The associated EPs are
$p(\x) = P(\gp[\x,1] \leq t_1,\gp[\x,2] \leq t_2)$,
$\x \in \mathcal{M}$.

For statistical generality, we will however keep the abstract notation
with $\T$ and general $\es$ \kc{Again, previous phrase doesn't parse.}
for most of the derivations. Note that assuming that $\gp$ has
continuous trajectories (almost surely) and $T$ to be closed, then
$\es$ can be cast as a Random Closed Set
\citep{Molchanov2005}. %In what follows, we will express various quantities related to $\es$ in terms of $\gp$'s moments both for general and orthant $T$'s.
Let us in particular consider a locally finite (Borel) measure $\mes$
on $D$ and investigate the probability distribution of $\mes(\es)$
through its moments.

@@

%That $\mu(\es)$ is a well-defined random variable follows from
\begin{propo}
	\label{propo1}
	%Under technical conditions to be precised, $\mu(\es)$ possesses moments or arbitrary order and these can be expressed for any $r\geq 1$ as
	For a measurable random field $\gp$ and a locally finite
        measure $\mes$ on $\mathcal{M}$, $\mes(\es)$ is a random
        variable and for any $r\geq 1$, \begin{equation*}
          \begin{split} \mathbb{E}[\mes(\es)^r]
            &=\int_{\mathcal{M}^{r}} \jointExcuProb \productMeasure ,
          \end{split} \end{equation*}

	where the product measure is denoted as
	$\nu^{\otimes}:=\bigotimes_{i=1}^r \nu$. 
	%Note that we will use boldface to denote concatenated variables. % TO RELOCATE
	Here $\gp$ is defined on $\mathcal{M}$, and for
	$\bm{u}=\left(u^{(1)}, ..., u^{(r)}\right)\in \mathcal{M}^r$, $\gp[\bm{u}]=\left(\gp[u^{(1)}], ...,
	\gp[u^{(r)}]\right)\in \mathbb{R}^{\no r}$.\kc{Is $\no r$ a
        product of the $\no$ from above? Is that obvious?}
	\medskip

	In the particular case where $\gp$ is a multivariate Gaussian random field 
	%and denoting the mean and covariance matrix of $\gp[\bm{u}]$ by $\meanUU\in\mathbb{R}^{p\times r}$ and $\covUU$, respectively, 
	we have 
	\begin{align*}
	\jointExcuProb = \mathcal{N}_{\no r}(T^r; \meanUU, \covUU),
	\end{align*}
	where $\mathcal{N}_{\no r}(\cdot ; \meanUU, \covUU)$ is the Gaussian measure on $\mathbb{R}^{\no r}$ with mean $\meanUU$ and covariance matrix $\covUU$, respectively defined blockwise by  
	%where both are defined blockwise by 
	\begin{align*}
	\meanUU&=\begin{pmatrix}\mu(u^{(1)})\\ \vdots\\ \mu(u^{(r)})\end{pmatrix}
	\in \mathbb{R}^{\no r}, \\
	\text{and } \covUU &= \begin{pmatrix}
	\cov(\gp[u^{(1)}], \gp[u^{(1)}]) & \dots & \cov(\gp[u^{(1)}],
	\gp[u^{(r)}])\\
	\vdots & & \vdots\\
	\cov(\gp[u^{(r)}], \gp[u^{(1)}]) & \dots & \cov(\gp[u^{(r)}],
	\gp[u^{(r)}])\\
	\end{pmatrix}\in \mathbb{R}^{pr\times pr},
	\end{align*}
	each of the $r\times r$ blocks of the latter matrix being itself a (cross-)covariance matrix of dimension $\no \times \no$. 
	%dimensional Gaussian random vector. 
	Assuming further that $\covUU$ is non-singular, the probability of interest can be formulated in terms of the $\no r$-dimensional Gaussian probability density function 
	$\varphi_{\no  r}(\cdot;~\meanUU, \covUU)$ as 
	%$\varphi_{\no \times r}$ as
	%$\bm{u}\in D^r$ such that $\gp[\bm{u}]$ be non-degenerate, the integrand can be expressed in terms of the $r \times \no$-dimensional Gaussian distribution, namely 
	%cumulative distribution function, namely 
	\begin{equation*}
	\begin{split}
	&\jointExcuProb
	=
	\int_{T^r} \varphi_{\no  r}\left(\bm{v};
	~\meanUU, \covUU\right) 
	\mathrm{d}\bm{v},
	\end{split}
	\end{equation*}
	In the particular orthant case with $\T=(-\infty, t_1] \times \dots \times (-\infty, t_{r}]$,
	the latter probability directly writes \kc{What does
          ``writes'' mean here? Connotes?} in terms of the multivariate Gaussian
	cumulative distribution, % with mean $\meanUU$ and covariance matrix $\covUU$,
	this time without requiring $\covUU$ %the latter 
	to be non-singular: 
	\begin{equation*}
	\begin{split}
	\jointExcuProb 
	&=
	\varPhi_{\no r}\left(\bm{t};~\meanUU, \covUU\right),
	\end{split}
	\end{equation*}
	where we have used the notations 
	$t=(t_1,\dots,t_{\no}%,..., t_1, ..., t_p
	)\in\mathbb{R}^{\no}$, $1_{r}=(1,\dots,1)\in \R^{r}$, and 
	$\bm{t}=1_{r}\otimes \bm{t}=(t_1,\dots,t_{\no},\dots,t_1,\dots,t_{\no}) 
	\in \R^{\no r}$.
\end{propo}

Proposition~\ref{propo1} enables one to calculate centered moments of $\mes(\es)$. 
The excursion measure variance $\emv = \operatorname{Var}[\mes(\es)]$
is probably the most relevant of these for our purposes. We
have 
\begin{equation*}
\begin{split}
\emv
&=\int_{\mathcal{M}^2} \mathbb{P}\left(
\gp[u]\in T, \gp[v]\in T \right) 
d\mes^{\otimes}(u, v)\\
&-\left( \int_{\mathcal{M}} \mathbb{P}\left(\gp[u]\in T\right) d\mes(u) \right)^2,
\end{split}
\end{equation*} 
which evaluates in the excursion/sojourn case where
$\T=(-\infty, t_1] \times \dots \times (-\infty, t_{\no}]$ to
\begin{equation*} \begin{split}
  \emv
%\operatorname{Var}[\mes(\es)]
&=\int_{\mathcal{M}^2} 
\varPhi_{2\no}
\left(
(\bt, \bt); \bmu((u,v)), 
K((u,v),(u,v))
\right) 
\
\mathrm{d}\mes^{\otimes} %\mes 
%\productMeasure 
(u,v)\\
&-\left( \int_{\mathcal{M}} \varPhi_{\no}\left(\bt;\bmu(u), K(u)\right) d\mes(u) \right)^2.
\end{split}
\end{equation*}
This quantity requires  an integral over $D^2$. In
contrast, the Integrated Bernoulli Variance (IBV) of \cite{Bect.etal}
involves solely an integral on $\mathcal{M}$ and can be expanded in
our present setting as follows 
\begin{equation}\label{IBV}
\begin{split}
\operatorname{IBV} %(\es) %;\mes)
&=\int_{\mathcal{M}}
\mathbb{P}\left(\gp[\uu]\in T\right)(1-\mathbb{P}\left(\gp[\uu]\in T\right))
d\mes(u) \\
&=\int_{\mathcal{M}}
\varPhi_{\no}\left(\bt;\bmu(\uu), K(\uu)\right) 
-\left(\varPhi_{\no}\left(\bt;\bmu(\uu), K(\uu)\right) \right)^2
\mathrm{d}\mes(u). 
\end{split}
\end{equation}

\subsection{Expected Integrated Bernoulli Variance and Sampling Designs}
\label{sec:EIBV}

Our goal is to construct informative sampling designs, given limited
resources \kc{information?}. In our domain, data is gathered by an AUV
that can measure temperature and salinity at chosen locations on a
defined waypoint graph. Such a discretized grid fosters reliable
navigation and mission performance. The waypoints further enable a
structure to the design problem in that, that sampling flexibility is
limited to operation on the graph. % This gives lots of sampling
% opportunities, and it is important to spend the sampling efforts
% wisely.
We focus on sampling designs that reduce the expected integrated
Bernoulli variance (EIBV). Note that this sampling can be done in a
static manner where the survey design is pre-planned, or in an
adaptive setting. For the latter case, data can be processed
sequentially and the updated model is then used to design and select
the next waypoint graph to visit, which in turn are used again to
update the model, and so on (Section \ref{sec:heuristics}). We next
describe the data modeling assumptions for one stage and develop
design criteria of semi-analytical forms relying on integrals of
multivariate Gaussian cumulative distribution functions; we will
appeal to some key intermediate results presented next.

>>>>

The remainder of this paper is organized as follows: 
%Section \ref{sec:bg} provides select background on ocean sampling. 
Section \ref{sec:ESEP} defines ESs, EPs, and the design criteria of IBV for
vector-valued GPs. Section \ref{sec:heuristics} builds on these
assumptions when deriving the sequential design criteria for adaptive
sampling. Section \ref{sec:simulations} discusses properties of the
methods in simulation studies. Section \ref{sec:case_study}
demonstrates the methodology used in field work characterizing a river
plume and finally, Section \ref{sec:concl_disc} contains a summary and
a discussion of future work.
%\input{section2}

\section{Quantifying uncertainty on Excursion Sets implicitly defined by Gaussian processes}
\label{sec:ESEP}

In section \ref{sec:bg_and_notation} we introduce the notation used in the rest of the paper, review (co-)kriging of multivariate gaussian random fields (GRFs) and present update formulae for cokriging.
Uncertainty quantification (UQ) techniques on excursion sets of GRFs are presented in \ref{sec:set_uq}, in particular the Integrated Bernoulli Variance (IBV) and the excursion measure variance (EMV) are defined and semi-analytical expressions easing their computation and optimization are derived.
Section \ref{sec:eibv} turns to the  effect of new observations on EMV and IBV. Specifically, the expected effect of the inclusion of a new set of observations on the EMV and IBV are considered and further semi-analytical expressions for the expected 
uncertainty reduction are derived. The resulting formulae form the backbone of the uncertainty reduction strategies presented in Section \ref{sec:heuristics}.
Finally, Section \ref{Sec:UnivarEx} illustrates the application of the above concepts on a bivariate example relevant for sampling in the temperature and salinity application.

\begin{comment}
\medskip

%Our objective is to leverage statistical tools onboard an autonomous robotic platform to characterize a river plume, focusing on spatial separation of cold freshwater from a river and warmer saline waters of a fjord.

The temperature and salinity field over the domain of interest will be modeled as a bivariate gaussian random field. The \textcolor{red}{complement of the} river plume may then be characterized as an \textit{excursion set} of the field, that is, as the region of the domain for which temperature and salinity lie above some specified threshold. Our goal is then to guide the data collection process so as to improve the characterization of this set.
The methods we develop are fairly general, and not limited to the special case of two-dimensional random fields. Hence, the rest of this section will consider generic excursion sets of gaussian random fields with an arbitrary number of output dimensions.

\begin{figure}[h!] \centering
  \includegraphics[width=0.99\textwidth]{Figures/example_excu_1.png}
  \caption{Realization of a bivariate GRF and excursion set above some threshold. Joint excursion in red, excursion of a single variable in white.}
\label{example_excu}
\end{figure}

Say we have an underlying phenomenon that is modeled as a $\no$-variate gaussian random field $\gp$ on some domain $\domain$, 
and we are interested in recovering the set of locations $\es$ in the domain for which the response variables (the components of $\gp$) lie in some set of specified values $\T\subset \mathbb{R}^{\no}$:
$$
\es:=\gp^{-1}(\T)=\{\x \in \mathcal{M}: \gp[\x] \in \T\}.
$$
Our goal here is to develop approaches to quantify and improve the
characterization of uncertainties on $\es$. \textcolor{red}{Alread mention how observations are included.}

If we assume that $\gp$ has
continuous trajectories (almost surely) and $T$ is closed, then
$\es$ becomes a Random Closed Set
\citep{Molchanov2005} and concepts from the theory of random sets will prove useful in characterizing the distribution of the volume of $\es$ under some measure defined on the domain.

\textcolor{red}{Is this really needed?} Note that while some aspects of the developed approaches do not call for a
specific form of $\T$, we will often stick for simplicity to the case
of orthants
($\T=(-\infty, t_1] \times \dots \times (-\infty, t_{\no}]$ where
$t_1,\dots, t_{\no} \in \R$) as this will allow efficient calculation
of several key quantities. Note that changing some $\leq$ inequalities
to $\geq$ ones would lead to immediate adaptations.

\end{comment}
\subsection{Background, Notation and Co-Kriging}
\label{sec:bg_and_notation}

Throughout the rest of the article, we denote by $\gp$ a vector-valued random field indexed by some arbitrary domain $\domain$, and assume values of the field at any fixed location $\x \in \domain$, denoted $\gp[\x]$, to be a $\no$-variate random vector ($\no\geq 2$). In our leading river plume characterization application, $\domain$ is a prescribed domain in the fjord of Trondheim (for the purpose of our AUV application, a discretization of a $2$-dimensional domain at fixed depth is considered), while $\no=2$ and the two responses of interest stand for temperature and salinity. A bivariate gaussian random field model is assumed for $\gp$, see the two first panels of Figure~\ref{example_excu} for an illustration of a realization of such a vector-valued model (RK NOTATION). The right panel of this figure represents a by-product of interest derived from these realizations, namely regions i) in red, where both temperature and salinity are high (i.e., exceeding respective thresholds), standing for ocean water ii) in blue, where both temperature and salinity are low, standing for riverine water, and iii) in white, where one variable is above and the other below their respective thresholds, standing for mixed waters.

\begin{figure}[h!] \centering
  \includegraphics[width=0.99\textwidth]{Figures/example_excu_1.png}
  \caption{Realization of a bivariate GRF and excursion set above some threshold. Joint excursion in red, excursion of a single variable in white.}
\label{example_excu}
\end{figure}

%The methods we develop are fairly general, and not limited to the special case of two-dimensional random fields. Hence, the rest of this section will consider generic excursion sets of gaussian random fields with an arbitrary number of output dimensions.
% Say we have an underlying phenomenon that is modeled as a $\no$-variate gaussian random field $\gp$ on some domain $\domain$, 
Coming back to the more general settings of a $\no$-variate random field, 
we are interested in recovering the set of locations $\es$ in the domain for which the response variables (the components of $\gp$) lie in some set of specified values $\T\subset \mathbb{R}^{\no}$, in other words \textit{the pre-image of $T$ by $\gp$}:
$$
\es:=\gp^{-1}(\T)=\{\x \in \mathcal{M}: \gp[\x] \in \T\}.
$$
%
%Our goal here is to develop approaches to quantify and improve the
%characterization of uncertainties on $\es$. \textcolor{red}{Alread mention how observations are included.}

If we assume that $\gp$ has
continuous trajectories (almost surely) and $T$ is closed, then
$\es$ becomes a Random Closed Set
\citep{Molchanov2005} and concepts from the theory of random sets will prove useful in characterizing the distribution of the volume of $\es$ under some measure defined on the domain.
%\textcolor{red}{Is this really needed?} 
Note that while some aspects of the developed approaches do not call for a
specific form of $\T$, we will often stick for simplicity to the case
of orthants
($\T=(-\infty, t_1] \times \dots \times (-\infty, t_{\no}]$ where
$t_1,\dots, t_{\no} \in \R$) as this will allow efficient calculation
of several key quantities. Note that changing some $\leq$ inequalities
to $\geq$ ones would lead to immediate adaptations.

@@

%\subsubsection{Notation}
%In order to work at this level of generality, some notational tricks are needed. Indeed, when considering a new observation, one shall in general specify which component of the field was observed and where it was observed. This leads us to introduce the concept of \textit{generalized location}.

Letting $\gp[\spatloc,\ell]$ denote the $\ell\text{-th}$ component of $\gp[\spatloc]$, $\spatloc\in \domain$ and $\ell \in \{1\dots,p\}$, we will call \textit{generalized location} the couple $x=(\spatloc,\ell)$. 
%Given such a generalized location notation $x$,  
The notation $\gp[x]$ will be used to denote $\gp[\spatloc,\ell]$ 
and
%
%This slight change of notation 
will allow us to think of $\gp$ as a scalar-valued Gaussian random field 
indexed by $D\times \{1\dots,p\}$, which will give the co-kriging equations a particularly simple form that parallels the one of univariate kriging. Due to the naturality of the concept of \textit{generalized location}, 
we will generally use the word \textit{location}, while \textit{spatial location} will be used to stress that we are talking about a point $\spatloc \in \domain$. In the following, the letter $x$ will be reserved for generalized locations, while the letters $\spatloc$ and $\ell$ will be used for spatial locations and response indices respectively.

It turns out that the inclusion of several observations at a time (a batch) may be handled by the same equations as the 
one for a single observation, provided some notation adjustments are made. This motivates the following.

Given a dataset consisting of  $q$ observations at spatial locations $\spatloc_i \in \domain$ and response indices $l_i \in \lbrace 1, ..., \no\rbrace$, $i=1, ..., q$, we can concatenate the information needed to specify the dataset by using the notation
\begin{align*}
\bm{x}:=(\bm{\spatloc}, \bm{\ell}):= (x_1,\dots, x_q),~\text{with }x_i=(\spatloc_i,\ell_i).
\end{align*}
In general, boldface letters will be used to denote concatenated quantities corresponding to batches of observations. 
We can then also compactly denote the values of the field at those different locations by
\begin{align*}
\gp[\bm{x}]:=
\left(\gp[\spatloc_1,\ell_1], ...,
\gp[\spatloc_q,\ell_{q}]\right) \in \mathbb{R}^{q}.
\end{align*}
%

Given a random field $\gp$ and (noisy) observations of some of its components at some points in the domain, one can predict the value of the field at some unobserved location $\spatloc\in \domain$ by using the conditional mean of $\gp[\spatloc]$, conditional on the data. This process is called (co-)kriging, and kriging equations precisely tell us how to compute conditional means and covariances conditional on an arbitrary dataset.
We will here present the most general form of cokriging, where one can include several observations (batch) at a time and 
observations at a given location $s \in \domain$ may only include a subset of the components of $\gp[\spatloc]\in\mathbb{R}^{\no}$ 
(heterotopic).

\subsubsection{Co-Kriging}
\label{sec:cokriging}
Assuming now that $n$ batches of observations are available, with respective sizes $q_1,\dots, q_n$, and that one wishes 
to predict $\gp[\bm{x}]$ for some batch of $q\geq 1$ generalized locations $\bm{x}=(\bm{\spatloc}, \bm{\ell})$, the (simple) 
cokriging mean would then amount to simple kriging with respect to a scalar-valued Gaussian random field indexed by 
$D\times \{1\dots,p\}$ and with covariance kernel $k(\bm{x}, \bm{x}')=K(\bm{\spatloc}, \bm{\spatloc}')_{\ell, \ell'}$, that is:
%
\begin{equation}
\mu_{[n]}(\bm{x})=\mu(\bm{x})+\lambda_{[n]}(\bm{x})^T (\mathbf{z}_{[n]}-\mu(\bm{x})),
\end{equation}
where $\mu$ is $Z$'s initial mean function, $\mathbf{z}_{[n]}$ stands for the ($\sum_{i=1}^n q_i$)-dimensional vector of 
observed responses of $Z$ at all considered generalized locations, and $\lambda_{[n]}(\bm{x})$ is a vector of weights 
equal to $k(\bm{x}_{[n]}, \bm{x}_{[n]})^{-1} k(\bm{x}_{[n]}, \bm{x})$ with $\bm{x}_{[n]}=(\bm{x}_1,\dots, \bm{x}_n)$, 
$k(\bm{x}_{[n]}, \bm{x}_{[n]})$ being assumed non-singular throughout the presentation. The co-kriging %(conditional) 
residual (cross-)covariance function (with respect to batches of generalized locations) can also be expressed in the same vein via
%
\begin{equation}
k_{[n]}(\bm{x},\bm{x}')=k(\bm{x},\bm{x}')-\lambda_{[n]}(\bm{x})^T k(\bm{x}_{[n]}, \bm{x}_{[n]}) \lambda_{{[n]}}(\bm{x}').
\end{equation}

\subsubsection{Co-kriging update formulae}

Let us now consider the case where co-kriging prediction of $Z$ was made with respect to $n$ batches $\bm{x}_i$ of generalized locations, concatenated again within
$\bm{x}_{[n]}=(\bm{x}_1,\dots, \bm{x}_n)$, and one wishes to update the prediction by incorporating a new vector of observations $\mathbf{z}_{n+1}$ measured at a batch of $q_{n+1} \geq 1$ generalized locations $\bm{x}_{n+1}$.
%It turns out that the concept of \textit{generalized location} makes the kriging formulae form-invariant across all dimensions. This allows us to directly adapt the 
Thanks to our representation of co-kriging in terms of simple kriging with respect to generalized locations, a strightforward adaptation of the batch-sequential kriging update formulae from \cite{Chevalier.etal2013a} delivers that
% 
\begin{equation}
\mu_{[n+1]}(\bm{x})=\mu_{[n]}(\bm{x})+\lambda_{[n+1,n+1]}(\bm{x})^T (\mathbf{z}_{n+1}-\mu(\bm{x}_{n+1})),
\end{equation}
where $\lambda_{[n+1,n+1]}(\bm{x})$ denotes the $q_{n+1}$-dimensional sub-vector extracted from
$\lambda_{[n+1]}(\bm{x})$ that corresponds to the kriging weigths associated with the last $q_{n+1}$ responses when 
predicting at $\bm{x}$ relying on all measurements until batch $(n+1)$.
%\text{th}$ batch.
%, i.e. those from the $(n+1)\text{th}$ batch of measurements conducted at $\bm{x}_{n+1}$.
Similarly, the updated co-kriging residual (cross-)covariance function then writes
\begin{equation}
k_{[n+1]}(\bm{x},\bm{x}')=k_{[n]}(\bm{x},\bm{x}')-\lambda_{[n+1,n+1]}(\bm{x})^T k_{[n]}(\bm{x}_{[n]}, \bm{x}_{[n]}) \lambda_{{[n+1,n+1]}}(\bm{x}').
\end{equation}
%\medskip
%
Let us remark that, as noted in \cite{Chevalier2015} in the case of scalar-valued fields, these update formulae naturally 
extend to Universal Kriging in second-order settings and apply without Gaussian assumption. We will now see how the latter formulae are instrumental in deriving semi-analytical formulae for stepwise uncertainty reduction criteria for 
vector-valued random fields.

\subsection{Uncertainty Quantification on Excursion Sets of multivariate Gaussian Random Fields}
\label{sec:set_uq}
We now introduce quantities that allow to quantify the uncertainty on the volume of the excursion set $\es$. Let $\mes$ be a 
(locally finite, Borel) measure  on $\domain$. We want to investigate the probability distribution 
of $\mes(\es)$ through its moments.
